Finding a closest point in a lattice of Voronoi's first kind

### 2013-12-18 paper.tex
minor
added derivation of Sellings formula

\usepackage{../bib/mathbf-abbrevs}

@@

\begin{lemma}\label{lem:decSellings}
Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbasis.  The following equalities hold:
\begin{enumerate}
\item  ${\displaystyle \|Bp\|^2 - \|B(p - I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i)}$, \label{eq:lem:decSellingsdec}
\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
\end{enumerate}
where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.
\end{lemma}
\begin{proof}
We give a proof for part~\ref{eq:lem:decSellingsdec}.  The proof for part~\ref{eq:lem:decSellingsinc} is similar.  From Selling's formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874},
\begin{align*}
\|Bp\|^2 &= \| \sum_{i=1}^{n+1} b_i p_i \|^2 \\
&= - \sum_{i=1}^{n+1}\sum_{j=i+1}^{n+1}q_{ij}(p_i - p_j)^2 \\
&= - \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}(p_i - p_j)^2, 
\end{align*}
where $q_{ij} = q_{ji} = b_i \cdot b_j$ are the Selling parameters.  Put $g = p - I_S$.  It follows that
\[
\|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
\]
where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
\[
d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
\]
If $i \in S$ and $j \notin S$ then
\[
d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
\]
while if $i \notin S$ and $j \in S$ then
\[
d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
\]
We thus have
\begin{align*}
\|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
\end{align*}
where the last line holds because $q_{ij} = q_{ji}$.
\end{proof}

@@

%\small
\bibliography{../bib/bib}

>>>>

\usepackage{mathbf-abbrevs}

@@

\begin{lemma}\label{lem:decSellings}
Let $p \in \reals^{n+1}$, let $S$ be a subset of the indices of $p$, and let $B$ be the matrix from~\eqref{eq:matrxBobtusebasis} with column given by the superbasis vectors $b_1,\dots,b_{n+1}$.  The following equalities hold:
\begin{enumerate}
\item  ${\displaystyle \|Bp\|^2 - \|B(p - I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i)}$, \label{eq:lem:decSellingsdec}
\item  ${\displaystyle \|Bp\|^2 - \|B(p + I_S)\|^2 = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)}$, \label{eq:lem:decSellingsinc}
\end{enumerate}
where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters.
\end{lemma}
\begin{proof}
We give a proof for part~\ref{eq:lem:decSellingsdec}.  The proof for part~\ref{eq:lem:decSellingsinc} is similar.  Observe that
\[
\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_i^2 =  \sum_{i=1}^{n+1}p_i^2\sum_{j=1}^{n+1} q_{ij} = 0
\]
and
\[
\sum_{i=1}^{n+1}\sum_{j=1}^{n+1} q_{ij} p_j^2 =  \sum_{j=1}^{n+1}p_j^2\sum_{i=1}^{n+1} q_{ij} = 0
\]
since $\sum_{i=1}^{n+1} q_{ij} = 0$ as a result of the superbasis condition~\eqref{eq:superbasecond}.  Now
\begin{align*}
\|Bp\|^2 &= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} p_i p_j \\
&= \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (\tfrac{1}{2}p_j^2 + p_i p_j -  \tfrac{1}{2}p_i^2) \\
&= -\frac{1}{2} \sum_{i=1}^{n+1} \sum_{j=1}^{n+1} q_{ij} (p_i - p_j)^2.
\end{align*}
This is closely related to what is known as Sellings formula~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  Put $g = p - I_S$.  It follows that
\[
\|Bp\|^2 - \|Bg\|^2 = \frac{1}{2} \sum_{i=1}^{n+1}\sum_{j=1}^{n+1}q_{ij}d_{ij},
\]
where $d_{ij} = (g_i - g_j)^2 - (p_i - p_j)^2$.  If both $i \notin S$ and $j \notin S$ then $p_i = g_i$, $p_j = g_j$ and so $d_{ij} = 0$.  Similarly if both $i \in S$ and $j \in S$ then 
\[
d_{ij} = (p_i-1 - p_j+1)^2 - (p_i - p_j)^2 = 0.
\]
If $i \in S$ and $j \notin S$ then
\[
d_{ij} = (p_i-1 - p_j)^2 - (p_i - p_j)^2= 1 + 2p_j - 2p_i,
\]
while if $i \notin S$ and $j \in S$ then
\[
d_{ij} = (p_i - p_j+1)^2 - (p_i - p_j)^2 = 1 + 2p_i - 2p_j.
\]
We thus have
\begin{align*}
\|Bp\|^2 - \|Bg\|^2 &= \frac{1}{2} \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i) + \frac{1}{2} \sum_{i \notin S}\sum_{j \in S}q_{ij}(1 + 2p_i - 2p_j) \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_j - 2p_i),
\end{align*}
where the last line holds because $q_{ij} = q_{ji}$.
\end{proof}

@@

%\small
\bibliography{bib}
